### 2022-11-12 paper_supplementary.Rmd

---
title: "Supplementary Figures: Quality control for the target decoy approach for peptide identification"
author: 
  Elke Debrie$^1$
  Milan Malfait$^1$
  Ralf Gabriels$^{2,3}$
  Arthur Declerq$^{2,3}$
  Adriaan Sticker$^{2,3}$
  Lennart Martens$^{2,3}$
  Lieven Clement$^{1,4}$
date: 
  $^1$ Department of Applied Mathematics, Computer Science and Statistics, Ghent University, Ghent, Belgium \newline
  $^2$ VIB-UGent Center for Medical Biotechnology, VIB, Zwijnaarde, Belgium \newline
  $^3$ Department of Biomolecular Medicine, Ghent University, Ghent, Belgium \newline
  $^4$ Corresponding Author$:$ lieven.clement@ugent.be
output: 
  bookdown::pdf_document2:
    keep_tex: yes
    latex_engine: xelatex
    citation_package: natbib
    toc: no
link-citations: yes
bibliography: [biblio.bib]
biblio-style: biochem
linkcolor: blue
urlcolor: blue
citecolor: blue
---

\setcitestyle{super,open={[},close={]}}
\def\citenumfont{}

(ref:Figure1)  Histogram and PP-plot for a concatenated search on a *Pyrococcus* run against a database of canonical *P. furiosus* sequences from Swiss-Prot using X!Tandem without refinement. Both the histogram and the P-P plot show no violation of the TDA assumptions.

```{r sFig1, echo=FALSE, fig.cap="(ref:Figure1)",out.width="99%"}
knitr::include_graphics("./figs/figTandemNoRefineSwissHistPP.png")
```

(ref:Figure2)  Histogram and PP-plot for a concatenated search on a *H. sapiens* run against a database of *H. Sapiens* sequences from UniProt using MS-GF+. Both the histogram and the P-P plot show no violation of the TDA assumptions.

```{r sFig2, echo=FALSE, fig.cap="(ref:Figure2)",out.width="99%"}
knitr::include_graphics("./figs/figHumanMsgfPlus.png")
```

(ref:Figure3)  Histogram and PP-plot for a concatenated search on an immunopeptidomics run using Andromeda. Both the histogram and the P-P plot show no violation of the TDA assumptions.

```{r sFig3, echo=FALSE, fig.cap="(ref:Figure3)",out.width="99%"}
knitr::include_graphics("./figs/figPeptidomics.png")
```

(ref:Figure4)  Histogram and PP-plot for rank 2 target and decoy PSM scores of a concatenated search on a H. sapiens run against a database of H. Sapiens sequences from UniProt using MS-GF+.

```{r sFig4, echo=FALSE, fig.cap="(ref:Figure4)",out.width="99%"}
knitr::include_graphics("./figs/figHumanMsgfPlusR2.png")
```