### 2022-11-13 paper_supplementary.Rmd
comply with achemso template of JPR + add myTemplate.tex

---
title: "Supplementary Figures: Quality control for the target decoy approach for peptide identification"
output: 
  bookdown::pdf_document2:
    template: "myTemp.tex"
    keep_tex: yes
    latex_engine: xelatex
    toc: no
    citation_package: natbib
    includes:
      in_header: authors.tex

documentclass: achemso
classoption: 
  - journal=jprobs
  - manuscript=article

bibliography: [biblio.bib]
linkcolor: blue
urlcolor: blue
citecolor: blue
---

```{r echo=FALSE, warning=FALSE}
if (!tinytex::check_installed("achemso")) tinytex::tlmgr_install("achemso")
if (!tinytex::check_installed("caption")) tinytex::tlmgr_install("caption")
if (!tinytex::check_installed("setspace")) tinytex::tlmgr_install("setspace")
```

(ref:Figure1)  Histogram and PP-plot for a concatenated search on a *Pyrococcus* run against a database of canonical *P. furiosus* sequences from Swiss-Prot using X!Tandem without refinement. Both the histogram and the P-P plot show no violation of the TDA assumptions.

```{r sFig1, echo=FALSE, fig.cap="(ref:Figure1)",out.width="99%"}
knitr::include_graphics("./figs/figTandemNoRefineSwissHistPP.png")
```

(ref:Figure2)  Histogram and PP-plot for a concatenated search on a *H. sapiens* run against a database of *H. Sapiens* sequences from UniProt using MS-GF+. Both the histogram and the P-P plot show no violation of the TDA assumptions.

```{r sFig2, echo=FALSE, fig.cap="(ref:Figure2)",out.width="99%"}
knitr::include_graphics("./figs/figHumanMsgfPlus.png")
```

(ref:Figure3)  Histogram and PP-plot for a concatenated search on an immunopeptidomics run using Andromeda. Both the histogram and the P-P plot show no violation of the TDA assumptions.

```{r sFig3, echo=FALSE, fig.cap="(ref:Figure3)",out.width="99%"}
knitr::include_graphics("./figs/figPeptidomics.png")
```

(ref:Figure4)  Histogram and PP-plot for rank 2 target and decoy PSM scores of a concatenated search on a H. sapiens run against a database of H. Sapiens sequences from UniProt using MS-GF+.

```{r sFig4, echo=FALSE, fig.cap="(ref:Figure4)",out.width="99%"}
knitr::include_graphics("./figs/figHumanMsgfPlusR2.png")
```