### 2022-11-13 authors.tex
Deploying to gh-pages from  @ cc10a4a1fd0243cdda341bab08a8ed48ef11c95f 🚀

\author{Elke Debrie}
\affiliation[Ghent University]
{Department of Applied Mathematics, Computer Science and Statistics, Ghent University, Ghent University, Ghent, Belgium}

\author{Milan Malfait}
\affiliation[Ghent University]
{Department of Applied Mathematics, Computer Science and Statistics, Ghent University, Ghent University, Ghent, Belgium}

\author{Ralf Gabriels}
\affiliation[VIB-UGent]
{VIB-UGent Center for Medical Biotechnology, VIB, Zwijnaarde, Belgium}
\alsoaffiliation[Ghent University]
{Department of Biomolecular Medicine, Ghent University, Ghent, Belgium}

\author{Arthur Declerq}
\affiliation[VIB-UGent]
{VIB-UGent Center for Medical Biotechnology, VIB, Zwijnaarde, Belgium}
\alsoaffiliation[Ghent University]
{Department of Biomolecular Medicine, Ghent University, Ghent, Belgium}

\author{Adriaan Sticker}
\affiliation[Ghent University]{Department of Applied Mathematics, Computer Science and Statistics, Ghent University, Ghent University, Ghent, Belgium}
\alsoaffiliation[VIB-UGent]
{VIB-UGent Center for Medical Biotechnology, VIB, Zwijnaarde, Belgium}
\alsoaffiliation[Ghent University]
{Department of Biomolecular Medicine, Ghent University, Ghent, Belgium}

\author{Lennart Martens}
\affiliation[VIB-UGent]
{VIB-UGent Center for Medical Biotechnology, VIB, Zwijnaarde, Belgium}
\alsoaffiliation[Ghent University]
{Department of Biomolecular Medicine, Ghent University, Ghent, Belgium}

\author{Lieven Clement}
\affiliation[Ghent University]{Department of Applied Mathematics, Computer Science and Statistics, Ghent University, Ghent University, Ghent, Belgium}
\email{lieven.clement@ugent.be}