### 2022-11-13 index.Rmd
update theme index.html

---
title: Quality control for the target decoy approach for peptide identification
output:
    html_document:
      code_download: false
      toc: false
      toc_float: false
      number_sections: false
      theme: flatly
      highlight: tango
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

***

This repository contains all necessary files to perform all data analyses and generate all figures for our paper: "Quality control for the target decoy approach for peptide identification"

- Generate all Figures: [paper_figures](./paper_figures.html)

- Document with Supplementary Figures: [supplementary_figures](./paper_supplementary.pdf)