### 2022-11-17 index.Rmd
add paper + license

---
title: Quality control for the target decoy approach for peptide identification
output:
    html_document:
      code_download: false
      toc: false
      toc_float: false
      number_sections: false
      theme: flatly
      highlight: tango
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

***

This repository contains all necessary files to perform all data analyses and generate all figures for our paper: "Quality control for the target decoy approach for peptide identification"

- Generate all Figures: [paper_figures](./paper_figures.html)

- Document with Supplementary Figures: [supplementary_figures](./paper_supplementary.pdf)

- [bioRxiv preprint](./paper_main.pdf)

---

Links:

- data on on [zenodo](https://doi.org/10.5281/zenodo.7308022)

- [bioRxiv](https://doi.org/10.1101/2022.11.17.516857)

---

License:

The website and code are licensed under [CC BY-SA 4.0](http://creativecommons.org/licenses/by-sa/4.0/)

The main paper is licensed under [CC-BY-NC-ND 4.0](https://creativecommons.org/licenses/by-nc-nd/4.0/)